---
title: 'Model evaluation: Correlation coefficients'
author: Anne Aulsebrook
output: 
  rmdformats::robobook:
    toc_depth: 2
    code_folding: hide
---

*Purpose*: This notebook is used to assess the correlation between the true and predicted time that individuals spent on a behaviour. For each of the 'best' models (assessed in "eval-generate-performance-metrics.Rmd"), we calculated the Spearman’s rank correlation coefficient for each behaviour. 

The following steps are included:

* Import predictions from each of the 'best' models (assessed in "eval-generate-performance-metrics.Rmd")
* Match predictions with bird identity
* Subset to focal birds
* Calculate Spearman's correlation coefficient
* Save counts and coefficients

*Date created*: May 10 2023

```{r Load packages}
library(dplyr)
library(data.table) # for fread
library(purrr) # for map_df
library(flextable)
library(tidyr) # for replace_na
library(here)
```

```{r Set flextable defaults}
set_flextable_defaults(
  theme_fun = theme_apa,
  font.size=10,
  padding = 0.25,
  font.color = "black",
  font.family = "Calibri"
)
```

```{r Create list of focal birds}
focal_birds <- c("1681",
                 "1331",
                 "1361",
                 "1333",
                 "1326",
                 "1301",
                 "1368",
                 "G20-0529-B6.5", "G20-0529", # recorded differently in different files
                 "952",
                 "1372")
```

```{r Create functions}
count_by_bird <- function(df, id, truth, estimate, model_group=NA){
    # Counts and summarises the number of true and estimated instances of each behaviour by ID
    # model_id is used as an identifier within the output
    if (is.na(model_group)) {
        model_id = "unnamed model"
        
        count_truth <- df %>%
        group_by(.data[[id]], .data[[truth]]) %>%
        summarise(n_truth=n()) %>%
        rename(behaviour=.data[[truth]],
               id=.data[[id]])
        
        count_pred <- df %>%
        group_by(.data[[id]], .data[[estimate]]) %>%
        summarise(n_pred=n()) %>%
        rename(behaviour=.data[[estimate]],
               id=.data[[id]])
        
        all_counts <- full_join(count_truth, count_pred, 
                                by=c('behaviour', 'id'))
        
        all_counts$n_pred <- replace_na(all_counts$n_pred, 0)
        all_counts$n_truth <- replace_na(all_counts$n_truth, 0)
        
        ind_totals <- all_counts %>% 
        group_by(id) %>% 
        summarise(n_truth_total = sum(n_truth, na.rm=T),
                  n_pred_total = sum(n_pred, na.rm=T))
        
        all_counts <- left_join(all_counts, ind_totals, by='id') %>%
            mutate(perc_truth=n_truth/n_truth_total*100,
                   perc_pred=n_pred/n_pred_total*100,
                   model_id=model_id)
    }
    
    if (!is.na(model_group)) {
        count_truth <- df %>%
        group_by(.data[[model_group]], .data[[id]], .data[[truth]]) %>%
        summarise(n_truth=n()) %>%
        rename(model_id=.data[[model_group]],
               behaviour=.data[[truth]],
               id=.data[[id]])
        
        count_pred <- df %>%
        group_by(.data[[model_group]], .data[[id]], .data[[estimate]]) %>%
        summarise(n_pred=n()) %>%
        rename(model_id=.data[[model_group]],
               behaviour=.data[[estimate]],
               id=.data[[id]])
        
        all_counts <- full_join(count_truth, count_pred, 
                                by=c('model_id', 'behaviour', 'id'))
        
        all_counts$n_pred <- replace_na(all_counts$n_pred, 0)
        all_counts$n_truth <- replace_na(all_counts$n_truth, 0)
        
        ind_totals <- all_counts %>% 
        group_by(model_id, id) %>% 
        summarise(n_truth_total = sum(n_truth, na.rm=T),
                  n_pred_total = sum(n_pred, na.rm=T))
        
        all_counts <- left_join(all_counts, ind_totals, by=c('model_id','id')) %>%
            mutate(perc_truth=n_truth/n_truth_total*100,
                   perc_pred=n_pred/n_pred_total*100)
    }
    
    return(all_counts)
}

reduce_beh <- function(beh) {
    new_beh <- case_when(beh=="aggressive posturing" ~ "courtship/territorial",
                         beh=="dynamic squatting" ~ "courtship/territorial",
                         beh=="static squatting" ~ "courtship/territorial",
                         beh=="copulation attempt" ~ "courtship/territorial",
                         beh=="mounting male" ~ "courtship/territorial",
                         beh=="vigilance" ~ "rest/vigilance", 
                         beh=="resting" ~ "rest/vigilance",
                         beh=="being mounted" ~ "courtship/territorial",
                         TRUE ~ beh)
    return(new_beh)
    }
```

# Random Forests

For the stratified (randSTRAT and timeSTRAT) models, it is best to use the predictions for the test set. 
For the LSIO models (also referred to as LSIO or LOIO), there were not enough individuals for a test set, so instead must use the cross-validation predictions.

```{r Load RF top model predictions}
rf_reduc_lsio <- 
    fread(here("outputs","eval","RF_best-CV-predictions_reduced-classes.csv")) %>%
    filter(grepl("LSIO", model_id, fixed = TRUE))

rf_all_lsio <- fread(here("outputs","eval","RF_best-CV-predictions_original-classes.csv")) %>%
    filter(grepl("LSIO", model_id, fixed = TRUE))
    
rf_reduc_strat <- fread(here("outputs","eval","best_RF_test-predictions_reduced-classes.csv")) %>%
    filter(!grepl("LSIO", model_id, fixed=TRUE))

rf_all_strat <- fread(here("outputs","eval","best_RF_test-predictions_original-classes.csv")) %>%
    filter(!grepl("LSIO", model_id, fixed=TRUE))
```

```{r Reduce behaviours for reduced class predictions}
rf_reduc_lsio$truth_beh <- reduce_beh(rf_reduc_lsio$truth_beh)
rf_reduc_lsio$pred_beh <- reduce_beh(rf_reduc_lsio$pred_beh)
rf_reduc_strat$truth_beh <- reduce_beh(rf_reduc_strat$truth_beh)
rf_reduc_strat$pred_beh <- reduce_beh(rf_reduc_strat$pred_beh)
```

```{r Split LSIO models into separate dataframes}
# To add bird info, need to separate the models; row numbers match up differently
rf_reduc_lsio <- rf_reduc_lsio %>%
    mutate(transitions = case_when(grepl("ToutAll", model_id, fixed=TRUE) ~ "excluded",
                                  TRUE ~ "included"))
rf_reduc_lsio_t <- rf_reduc_lsio %>% filter(transitions=="included")
rf_reduc_lsio_nt <- rf_reduc_lsio %>% filter(transitions=="excluded")

rf_all_lsio <- rf_all_lsio %>%
    mutate(transitions = case_when(grepl("ToutAll", model_id, fixed=TRUE) ~ "excluded",
                                  TRUE ~ "included"))
rf_all_lsio_t <- rf_all_lsio %>% filter(transitions=="included")
rf_all_lsio_nt <- rf_all_lsio %>% filter(transitions=="excluded")
```

```{r Add bird info to lsio predictions}
# Note: bird info is already there for strat predictions
windowed_dat <- fread(here("data","windowed","windowed-data.csv"))
windowed_dat_nt <- windowed_dat %>% filter(transition==FALSE)
windowed_dat$.row <- seq(1, nrow(windowed_dat), 1) # add row number
windowed_dat_nt$.row <- seq(1, nrow(windowed_dat_nt), 1) # add row number

# Subset columns
windowed_sub <- windowed_dat %>% select(recording_id:majority_behaviour, .row)
windowed_sub_nt <- windowed_dat_nt %>% select(recording_id:majority_behaviour, .row)

# Join with bird info
rf_reduc_lsio_t <- full_join(windowed_sub, rf_reduc_lsio_t, by=".row")
rf_reduc_lsio_nt <- full_join(windowed_sub_nt, rf_reduc_lsio_nt, by=".row")
rf_all_lsio_t <- full_join(windowed_sub, rf_all_lsio_t, by=".row")
rf_all_lsio_nt <- full_join(windowed_sub_nt, rf_all_lsio_nt, by=".row")
```

```{r Clean redundant RF dataframes}
rm(windowed_dat_nt, rf_reduc_lsio, rf_all_lsio)
```

```{r Combine all predictions and add unique ID for model}
rf_preds <- bind_rows(lst(rf_all_lsio_nt,
                      rf_all_lsio_t,
                      rf_reduc_lsio_nt,
                      rf_reduc_lsio_t,
                      rf_all_strat,
                      rf_reduc_strat),
                      .id="model_spec")

rf_preds$model_id <- paste(rf_preds$model_spec, rf_preds$model_id)
```

```{r Count RF true and predicted behaviours}
rf_counts <- count_by_bird(rf_preds, 
                      id='ruff_id', 
                      truth='truth_beh', 
                      estimate='pred_beh',
                      model_group='model_id')
```

```{r Subset to RF focal birds and calculate correlation coefficient}
rf_counts_focal <- rf_counts %>% filter(id %in% focal_birds)
    
rf_cors <- rf_counts_focal |>
    split(f = list(rf_counts_focal$model_id, rf_counts_focal$behaviour), 
          drop = TRUE) |>
    map(function(x){
        result <- cor.test(x$perc_truth, x$perc_pred, method = "spearman", exact=FALSE)
        return(data.frame(cor_spearman = result$estimate, p = result$p.value))
    }) |> bind_rows(.id = "id") |>
    mutate(behaviour = gsub(".*[.]", "", id),
           model_id = gsub("[.].*", "", id)) |> 
    select(model_id, behaviour, cor_spearman, p) |>
    arrange(model_id)

```

# Hidden Markov Models

```{r Load and combine HMM predictions}
hmm_reduc_lsio <- fread(here("outputs","eval","HMM_CV_predictions_reduced-classes.csv")) %>%
    filter(split=="LSIO")
hmm_all_lsio <- fread(here("outputs","eval","HMM_CV_predictions_original-classes.csv")) %>%
    filter(split=="LSIO")

hmm_reduc_strat <- fread(here("outputs","eval","HMM_test_predictions_reduced-classes.csv")) %>%
    filter(split=="timesplit")
hmm_all_strat <- fread(here("outputs","eval","HMM_test_predictions_original-classes.csv")) %>%
    filter(split=="timesplit")

hmm_preds <- bind_rows(lst(hmm_reduc_lsio,
                      hmm_all_lsio,
                      hmm_reduc_strat,
                      hmm_all_strat),
                      .id="model_id")
```

```{r Add bird ID to HMM predictions}
hmm_preds <- hmm_preds %>%
    mutate(ruff_id=sub("_.*", "", recording_id))
```

```{r Count HMM true and predicted behaviours}
hmm_counts <- count_by_bird(hmm_preds, 
                      id='ruff_id', 
                      truth='truth_beh', 
                      estimate='pred_beh',
                      model_group='model_id')
```

```{r Subset to HMM focal birds and calculate correlation coefficient}
hmm_counts_focal <- hmm_counts %>% filter(id %in% focal_birds)

hmm_cors <- hmm_counts_focal |>
    split(f = list(hmm_counts_focal$model_id, hmm_counts_focal$behaviour), 
          drop = TRUE) |>
    map(function(x){
        result <- cor.test(x$perc_truth, x$perc_pred, method = "spearman", exact=FALSE)
        return(data.frame(cor_spearman = result$estimate, p = result$p.value))
    }) |> bind_rows(.id = "id") |>
    mutate(behaviour = gsub(".*[.]", "", id),
           model_id = gsub("[.].*", "", id)) |> 
    select(model_id, behaviour, cor_spearman, p) |>
    arrange(model_id)
```

# Neural networks

```{r Load and combine NN predictions}
nn_reduc_lsio <- fread(here("outputs","eval","NN_CV_predictions_reduced-classes.csv"))

nn_all_lsio <- fread(here("outputs","eval","NN_CV_predictions_original-classes.csv"))

nn_preds <- bind_rows(lst(nn_reduc_lsio,
                      nn_all_lsio),
                      .id="model_id")
```

```{r Add bird ID to NN predictions}
segment_ids <- windowed_dat %>%
    group_by(segment_id) %>%
    summarise(ruff_id=first(ruff_id),
              recording_id=first(recording_id))

nn_preds <- left_join(nn_preds, 
                      segment_ids,
                      by='segment_id')
```

```{r Count NN true and predicted behaviours}
nn_counts <- count_by_bird(nn_preds, 
                      id='ruff_id', 
                      truth='truth_beh', 
                      estimate='pred_beh',
                      model_group='model_id')
```

```{r Subset to NN focal birds and calculate correlation coefficient}
nn_counts_focal <- nn_counts %>% filter(id %in% focal_birds)

nn_cors <- nn_counts_focal |>
    split(f = list(nn_counts_focal$model_id, nn_counts_focal$behaviour), 
          drop = TRUE) |>
    map(function(x){
        result <- cor.test(x$perc_truth, x$perc_pred, method = "spearman", exact=FALSE)
        return(data.frame(cor_spearman = result$estimate, p = result$p.value))
    }) |> bind_rows(.id = "id") |>
    mutate(behaviour = gsub(".*[.]", "", id),
           model_id = gsub("[.].*", "", id)) |> 
    select(model_id, behaviour, cor_spearman, p) |>
    arrange(model_id)
```

# Tidy and export summarised data

```{r Tidy summaries}
join_counts <- rbind(hmm_counts, rf_counts, nn_counts) %>%
    mutate(model_type=fcase(grepl("rf",model_id,fixed=TRUE), "RF",
                                grepl("hmm", model_id, fixed=TRUE), "HMM",
                                grepl("nn", model_id, fixed=TRUE), "NN"),
           split_type=fcase(grepl("time", model_id, fixed=TRUE), "timeSTRAT",
                            grepl("lsio", model_id, fixed=TRUE), "LSIO",
                            grepl("hmm", model_id, fixed=TRUE) & grepl("strat", model_id, fixed=TRUE), "timeSTRAT",
                            grepl("strat", model_id, fixed=TRUE), "randSTRAT"),
           transitions_in_test=fcase(model_type=="HMM", as.character(NA),
                                     model_type=="NN", as.character(NA),
                                     grepl("ToutAll",model_id,fixed=TRUE), "no",
                                     model_type=="RF", "yes"),
           classes=fcase(grepl("reduc", model_id, fixed=TRUE), "reduced_classes",
                         grepl("all", model_id, fixed=TRUE), "original_classes")) %>%
    rename(ruff_id=id)

join_cors <- rbind(hmm_cors, rf_cors, nn_cors) %>%
    mutate(model_type=fcase(grepl("rf",model_id,fixed=TRUE), "RF",
                                grepl("hmm", model_id, fixed=TRUE), "HMM",
                                grepl("nn", model_id, fixed=TRUE), "NN"),
           split_type=fcase(grepl("time", model_id, fixed=TRUE), "timeSTRAT",
                            grepl("lsio", model_id, fixed=TRUE), "LSIO",
                            grepl("hmm", model_id, fixed=TRUE) & grepl("strat", model_id, fixed=TRUE), "timeSTRAT",
                            grepl("strat", model_id, fixed=TRUE), "randSTRAT"),
           transitions_in_test=fcase(model_type=="HMM", as.character(NA),
                                     model_type=="NN", as.character(NA),
                                     grepl("ToutAll",model_id,fixed=TRUE), "no",
                                     model_type=="RF", "yes"),
           classes=fcase(grepl("reduc", model_id, fixed=TRUE), "reduced_classes",
                         grepl("all", model_id, fixed=TRUE), "original_classes"))
```

```{r Write files}
write.csv(join_counts, 
         here("outputs","eval","true-vs-predicted_ind-beh-counts.csv"),
          row.names=FALSE)

write.csv(join_cors, 
         here("outputs","eval","correlation-coefficients.csv"),
         row.names=FALSE)
```
