---
title: "Creating tables and figures for publication"
authors: Anne Aulsebrook
output: 
  rmdformats::robobook:
    toc_depth: 2
    code_folding: hide
---

*Purpose*: This notebook is used to create the tables and figures included in the manuscript.

*Date created*: May 10 2023

```{r Set up packages and workspace}
library(data.table) # for fread
library(dplyr)
library(tidyr) # for pivot_wider
library(glue)
require(flextable) # for nice table outputs
library(officer) # for fp_border
library(stringr)
library(gdtools)
library(tibble)
library(ggplot2)
library(viridis) # for colour scaling
library(gridExtra)
library(here)
input_dir <- here("outputs","eval")
output_dir <- here("outputs","tables-figures")
options(huxtable.knit_print_df = TRUE)
options(scipen=100)
```

```{r Set number of digits}
digs = 2
```

```{r Set formatting parameters}
set_flextable_defaults(
  theme_fun = theme_apa,
  font.size=10,
  padding = 0.25,
  font.color = "black",
  font.family = "Calibri",
  big.mark = " ",
  na_str = " ",
  nan_str = " "
)

port_width <- 16
land_width <- 24
```

```{r Create list of focal birds}
focal_birds <- c("1681",
                 "1331",
                 "1361",
                 "1333",
                 "1326",
                 "1301",
                 "1368",
                 "G20-0529-B6.5", "G20-0529", # recorded differently in different files
                 "952",
                 "1372")
```

```{r Create functions for formatting}
rnd <- function(x){
    format(round(x, digs), nsmall = digs)
}

format_mean_sd <- function(mean, sd){
    val_with_sd <- case_when(is.na(sd) ~ rnd(mean),
                             TRUE ~ paste(rnd(mean), "±", rnd(sd)))
    return(val_with_sd)
}

rename_behaviours <- function(behs) {
    new_behs <- behs %>%
        str_replace("flying", "Fly") %>%
        str_replace("foraging or drinking", "Forage or drink") %>%
        str_replace("preening", "Preen") %>%
        str_replace("resting", "Rest") %>%
        str_replace("vigilance", "Vigilant") %>%
        str_replace("walking or running", "Walk or run") %>%
        str_replace("aggressive posturing", "Forward aggressive") %>%
        str_replace("dynamic squatting", "Dynamic squat") %>%
        str_replace("static squatting", "Static squat") %>%
        str_replace("being mounted", "Being mounted") %>%
        str_replace("copulation attempt", "Copulation attempt") %>%
        str_replace("mounting male", "Mounting male") %>%
        str_replace("other", "Other") %>%
        str_replace("courtship/territorial", "Mating behaviours") %>%
        str_replace("rest/Vigilant","Rest/vigilant")
    return(new_behs)
}

order_behaviours <- function(behs) {
    ordered_behs <- ordered(behs, 
                            levels = c("Fly",
                                              "Forage or drink",
                                              "Preen",
                                              "Rest",
                                              "Vigilant",
                                              "Walk or run",
                                              "Forward aggressive",
                                              "Dynamic squat",
                                              "Static squat",
                                              "Copulation attempt",
                                              "Mounting male",
                                              "Being mounted",
                                              "Mating behaviours",
                                              "Rest/vigilant",
                                              "Other"))
    return(ordered_behs)
}

rename_ids <- function(ids) {
    new_ids <- ids %>%
        str_replace("1681", "I1") %>% # orange-faced sweeney
        str_replace("1331", "I2") %>% # tiger white-bill
        str_replace("1361", "I3") %>% # foxy dalmation
        str_replace("1333", "I4") %>% # grey tabby
        str_replace("1326", "S1") %>% # snow grey
        str_replace("1301", "S2") %>% # lassie
        str_replace("1368", "S3") %>% # bald eagle
        str_replace("G20-0529-B6.5", "S4") %>% # ginger tabby
        str_replace("G20-0529", "S4") %>% # ginger tabby
        str_replace("952", "F1") %>% # pink faeder
        str_replace("1372", "F2") %>% # yellow faeder
        str_replace("G20-0055-B6.5", "S5") %>% # mustard-white
        str_replace("1399", "I5") %>% # dark dalmation
        str_replace("7-04-105", "I6") %>% # dark brown-scarf
        str_replace("G20-0071-B6.5", "S6") %>% # pale-head tiger
        str_replace("G20-0059-B6.5", "S7") # pale-head dalmation
    return(new_ids)
}

order_ids <- function(ids) {
    ordered_ids <- ordered(ids, 
                            levels = c("I1","I2","I3","I4",
                                       "S1","S2","S3", "S4",
                                       "F1","F2",
                                       "I5","I6","S5","S6","S7")
    )
    return(ordered_ids)
}
```

# Behaviour summary (Figure 2)

```{r Import windowed and behavioural data}
windowed_dat <- fread(here("data","windowed","windowed-data.csv"))
beh_full <- fread(here("data","clean", "ruff_behaviours_adjusted_for_drift.csv"))
```

```{r Extract key information about behaviours}
beh_counts <- windowed_dat %>%
    group_by(majority_behaviour) %>%
    rename(Behaviour = majority_behaviour) %>%
    summarise(total_dur=n(),
              `% transitions`=sum(transition==TRUE)/total_dur*100) %>%
    select(!total_dur)

beh_duration <- beh_full %>%
    group_by(behaviour) %>%
    rename(Behaviour = behaviour) %>%
    summarise(`total duration (s)`=sum(duration_secs),
              `mean bout (s)` = mean(duration_secs))

beh_counts <- left_join(beh_counts, beh_duration, by="Behaviour")
```

```{r Create behaviour table}
beh_counts$Behaviour <- rename_behaviours(beh_counts$Behaviour)
beh_counts$Behaviour <- order_behaviours(beh_counts$Behaviour)
beh_counts <- beh_counts %>% arrange(Behaviour)
beh_counts$`% transitions` <- rnd(beh_counts$`% transitions`)


beh_counts <- beh_counts %>%
    mutate(Description = case_when(Behaviour=="Fly" ~ "Flying",
                                    Behaviour=="Forage or drink" ~ "Handling food, drinking, or lowering head to search for food",
                                    Behaviour=="Preen" ~ "Preening feathers using bill",
                                    Behaviour=="Rest" ~ "Sitting or standing with eyes closing or bill tucked into feathers",
                                    Behaviour=="Vigilant" ~ "Sitting or standing with head upright, looking around",
                                    Behaviour=="Walk or run" ~ "Walking or running in an upright posture",
                                    Behaviour=="Forward aggressive" ~ "Standing with body angled forwards and bill horizontal. Can be accompanied by bill thrusting, stamping, tail trembling and/or charging",
                                    Behaviour=="Dynamic squat" ~ "Shuffling while in a horizontal posture, with bill pointed vertically at the ground",
                                    Behaviour=="Static squat" ~ "Standing or sitting still in a tense horizontal posture, with bill pointed at the ground",
                                    Behaviour=="Copulation attempt" ~ "Mounting female and attempting to copulate, bending tail downwards",
                                    Behaviour=="Mounting male" ~ "Mounting another male. Often includes shaking the tail from side to side",
                                    Behaviour=="Being mounted" ~ "Being mounted by another male",
                                   Behaviour=="Other" ~ "All other behaviours",
                                    TRUE ~ "error")) %>%
    relocate(Description, .after = Behaviour) %>%
    relocate(`% transitions`, .after = `mean bout (s)`)

beh_counts
```

```{r Make behaviour summary flextable, width = 50}
beh_col_width <- 3.5
desc_col_width <- 6

beh_flex <- beh_counts %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    colformat_double(j=3,
                     digits=0) %>%
    align(j=3:5, align="right", part="all") %>%
    valign(valign="top", part="body") %>%
    line_spacing(space = 1.5, part = "all") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(port_width-beh_col_width-desc_col_width)/(ncol(beh_counts)-2), unit="cm") %>%
    width(j=1, beh_col_width, "cm") %>%
    width(j=2, desc_col_width, "cm")

beh_flex
```

```{r Save behaviour table}
save_as_docx(beh_flex, path = here(output_dir, "behaviour_summary.docx"), 
             align = "left")
```

## Behaviour summary by individual (Table A?)

```{r Behaviour counts by individual}
beh_ind <- beh_full %>%
    group_by(behaviour, ruff_id) %>%
    rename(Behaviour = behaviour) %>%
    summarise(`total duration (s)`=sum(duration_secs),
              `mean duration (s)` = mean(duration_secs))

beh_ind$ruff_id <- rename_ids(beh_ind$ruff_id)
beh_ind$ruff_id <- order_ids(beh_ind$ruff_id)
beh_ind <- beh_ind %>% arrange(ruff_id)
beh_ind$Behaviour <- rename_behaviours(beh_ind$Behaviour)
beh_ind$Behaviour <- order_behaviours(beh_ind$Behaviour)
```

```{r Format individual behaviour duration}
beh_ind_formatted <- beh_ind %>%
    ungroup() %>%
    complete(Behaviour, ruff_id, fill = list(`total duration (s)` = 0,
                                             `mean duration (s)` = 0)) |>
    mutate(dur_combined=paste0(round(`total duration (s)`, 0),
                               "\n(", 
                               sprintf(round(`mean duration (s)`, 2), fmt = '%#.2f'), 
                               ")")) %>%
    select(Behaviour, ruff_id, dur_combined)

beh_ind_wide <- beh_ind_formatted %>%
    pivot_wider(.,
                names_from = ruff_id,
                values_from = `dur_combined`)

beh_ind_wide$Behaviour <- rename_behaviours(beh_ind_wide$Behaviour)
beh_ind_wide$Behaviour <- order_behaviours(beh_ind_wide$Behaviour)
beh_ind_wide <- beh_ind_wide %>% 
    arrange(Behaviour)
beh_ind_wide
```
```{r Create flextable for focal behaviour duration}
beh_col_width <- 3.5

beh_focal_flex <- beh_ind_wide %>%
    select(!(I5:S7)) %>%
    filter(Behaviour!="Mating behaviours" & Behaviour!="Rest/vigilant") %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    valign(valign="top", part="body") %>%
    line_spacing(space = 1.15, part = "all") %>%
    padding(padding.bottom = 10, part = "body") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(port_width-beh_col_width)/(ncol(beh_ind_wide)-6), unit="cm") %>%
    width(j=1, beh_col_width, "cm")

beh_focal_flex
```
```{r Create flextable for rare bird behaviour duration}
beh_col_width <- 3.5

beh_rare_flex <- beh_ind_wide %>%
    select(!(I1:F2)) %>%
    filter(Behaviour!="Mating behaviours" & Behaviour!="Rest/vigilant") %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    valign(valign="top", part="body") %>%
    line_spacing(space = 1.15, part = "all") %>%
    padding(padding.bottom = 10, part = "body") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(port_width-beh_col_width)/(ncol(beh_ind_wide)-6), unit="cm") %>%
    width(j=1, beh_col_width, "cm")

beh_rare_flex
```

```{r Save individual counts to get sense of size}
save_as_docx(`focal males`=beh_focal_flex,
             `other males`=beh_rare_flex,
             path = here(output_dir, "behaviour_ind_summary_duration.docx"), 
             align = "left")
```

# Acc recording summary (Table A?)

```{r Import and reformat recording information}
rec_info <- read.csv(here("data","raw","logger_deployment_notes.csv"))

# Datetimes entered into Excel spreadsheets can be reliable, so reextract release datetime
rec_info <- rec_info %>%
    mutate(release_dt=paste(capture_date, release_time))

rec_info <- rec_info %>%
    rename(ruff_id=ruff_id_number) %>%
    select(ruff_id, morph, release_dt, recapture_disturbance_start_dt) %>%
    mutate(start=as.POSIXct(release_dt, format="%d/%m/%Y %H:%M:%S"),
           end=as.POSIXct(recapture_disturbance_start_dt, format="%d/%m/%Y %H:%M")) %>%
    mutate(duration=as.numeric(difftime(end, start, units="days")),
           duration=format(round(duration, 1), nsmall = 1),
           start=format(start, format='%d.%m.%Y'),
           end=format(end, format='%d.%m.%Y'))
    
rec_info_wide <- rec_info %>%
    group_by(ruff_id, morph) %>%
    arrange(desc(start)) %>%
    slice(seq_len(3)) %>%
    mutate(recording=paste0("recording", row_number())) %>%
    select(-release_dt, -recapture_disturbance_start_dt) %>%
    pivot_wider(id_cols=c(ruff_id, morph),
                names_from=recording,
                values_from=c(start, end, duration)) %>%
    select(ruff_id, morph,
           start_recording1, duration_recording1,
           start_recording2, duration_recording2,
           start_recording3, duration_recording3)
```

```{r Make recording summary flextable}
first_col_width=4
second_col_width=2.6

rec_info_flex <- 
    # ---- Basic formatting ----
    rec_info_wide  %>%
    flextable() %>%
    valign(valign="center") %>%
    align(align = "left", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    font(fontname="Calibri", part="all") %>%
    colformat_datetime(na_str="NA") %>%
    colformat_num(na_str="NA") %>%
    # ---- Rename and format headings ----
    set_header_labels(ruff_id="Ruff ID",
                      morph="Morph",
                      start_recording1="Start",
                      duration_recording1="Duration (days)",
                      start_recording2="Start",
                      duration_recording2="Duration (days)",
                      start_recording3="Start",
                      duration_recording3="Duration (days)") %>%
    add_header_row(values = c("","Recording 1", "Recording 2", "Recording 3"),
                   colwidths = c(2, 2, 2, 2)) %>%
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(land_width-first_col_width-second_col_width)/(ncol(rec_info_wide)-2), unit="cm") %>%
    width(j=1, first_col_width, "cm") %>%
    width(j=2, second_col_width, "cm")

rec_info_flex
```

```{r Save acc recording table}
save_as_docx(rec_info_flex, path = here(output_dir, "acc_recording_summary.docx"), 
             align = "left")
```


# Macro-averaged model metrics (Table 2)

```{r Gather and combined macro-averaged metrics}
macro_og <- read.csv(here(input_dir, "top_macro-metrics_original-classes.csv"))
macro_reduc <- read.csv(here(input_dir, "top_macro-metrics_reduced-classes.csv"))

macro_og <- macro_og %>%
    mutate(classes="original_classes")
macro_reduc <- macro_reduc %>%
    mutate(classes="reduced_classes")

macro <- full_join(macro_og, macro_reduc)
```

```{r Summarise macro metrics}
macro_summary <- macro %>%
    mutate(split_val = paste(split_type, validation_type),
           all_prec=format_mean_sd(precision, precision_sd),
           all_recall=format_mean_sd(recall, recall_sd),
           all_fmeas=format_mean_sd(f_meas, f_meas_sd),
           mat_prec=format_mean_sd(precision_mat, precision_sd_mat),
           mat_recall=format_mean_sd(recall_mat, recall_sd_mat),
           mat_fmeas=format_mean_sd(f_meas_mat, f_meas_sd_mat)) %>%
    select(model_type, model_id, split_type, split_val, transitions_in_test, classes, all_prec:mat_fmeas)
```

```{r Macro table}
# ---- Restructure table ----
macro_table <- macro_summary %>%
    filter(transitions_in_test=="yes" | is.na(transitions_in_test)) %>%
    filter(split_val!="randstrat CV" & split_val!= "timesplit CV") %>%
    select(model_type, split_type, classes:mat_fmeas) %>%
    mutate(split_type = str_replace(split_type, "randstrat", "randSTRAT"),
           split_type = str_replace(split_type, "timesplit", "timeSTRAT"),
           classes = str_replace(classes, "original_classes", "Original classes"),
           classes = str_replace(classes, "reduced_classes", "Reduced classes"))

# ---- Set factor order ----
macro_table$model_type <- ordered(macro_table$model_type, levels = c("RF", "HMM", "NN"))
macro_table$split_type <- ordered(macro_table$split_type, levels = c("randSTRAT", "timeSTRAT", "LSIO"))

# ---- Arrange table ----
macro_table <- macro_table %>%
    arrange(classes, model_type, split_type) 

macro_table
```


```{r Macro flextable}
# ~~~~~ FORMAT MACRO TABLE -----------------------------------------------------
first_col_width <- 0.1

macro_table_flex <- 
    # ---- Basic formatting ----
    macro_table %>%
    # select(-all_fmeas, -mat_fmeas) %>%
    # filter(model_type=="RF" | split_type=="LSIO") %>%
    as_grouped_data(groups="classes") %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    add_header_row(values = c("","All behaviours", "Mating behaviours"),
                   colwidths = c(3, 3, 3)) %>%
    bold(part="header") %>%
    font(fontname="Calibri", part="all") %>%
    width(width=(port_width-first_col_width)/ncol(macro_table), unit="cm") %>%
    width(j=1, first_col_width, "cm") %>%
    # ---- Rename and format headings ----
    set_header_labels(classes="") %>%
    bold(i = ~ !is.na(classes), part="body") %>%
    set_header_labels(values = list(model_type = "Model type",
                                    split_type = "Split type",
                                    all_prec = "Precision",
                                    all_recall = "Recall",
                                    all_fmeas = "F1-score",
                                    mat_prec = "Precision",
                                    mat_recall = "Recall",
                                    mat_fmeas = "F1-score")) %>%
    merge_at(i=1, j=1:9, part="body") %>% # Original classes
    merge_at(i=8, j=1:9, part="body") %>% # Reduced classes
    # ---- Format 'model type' column ----
    merge_v(j = "model_type", target = "model_type", part = "body", combine = FALSE) %>%
    valign(valign="top", part="body") %>% # align
    # ---- Add borders ----
    hline(i=4, j=2:9, part="body") %>%
    hline(i=6, j=2:9, part="body") %>%
    hline(i=11, j=2:9, part="body") %>%
    hline(i=13, j=2:9, part="body")

macro_table_flex
```

```{r Save macro metrics table}
save_as_docx(macro_table_flex, path = here(output_dir, "macro_metrics_table.docx"), 
             align = "left")
```

# Class metrics (Table 3)

```{r Gather and combined class metrics}
class_og <- read.csv(here(input_dir, "top_class-metrics_original-classes.csv"))
class_reduc <- read.csv(here(input_dir, "top_class-metrics_reduced-classes.csv"))

class_og <- class_og %>%
    mutate(classes="original_classes")
class_reduc <- class_reduc %>%
    mutate(classes="reduced_classes")

class <- full_join(class_og, class_reduc)
```

```{r Summarise class metrics}
class_summary <- class %>%
    mutate(split_val = paste(split_type, validation_type),
           prec=format_mean_sd(precision, precision_sd),
           rec=format_mean_sd(recall, recall_sd),
           fmeas=format_mean_sd(f_meas, f_meas_sd)) %>%
    filter(classes=="original_classes"|behaviour=="courtship/territorial"|behaviour=="rest/vigilance") %>%
    select(model_type, model_id, split_type, split_val, transitions_in_test, behaviour, classes, prec:fmeas)
```

```{r Order factors for class metrics}
class_summary$behaviour <- rename_behaviours(class_summary$behaviour)
class_summary$behaviour <- order_behaviours(class_summary$behaviour)

class_summary$model_type <- ordered(class_summary$model_type, levels = c("RF", "HMM", "NN"))

class_summary$split_type <- ordered(class_summary$split_type, 
                                    levels = c("randstrat", "timesplit", "LSIO"))

class_summary$model_split_val <- paste(class_summary$model_type, class_summary$split_val)
```

```{r Class table with model columns}
# ---- Restructure table ----
class_table <- class_summary %>%
    filter(transitions_in_test=="yes" | is.na(transitions_in_test)) %>%
    filter(model_split_val!="RF randstrat CV" & 
               model_split_val!= "RF timesplit CV" &
               model_split_val!= "HMM timesplit CV") %>%
    select(model_type, split_type, behaviour:fmeas) %>%
    mutate(split_type = str_replace(split_type, "randstrat", "randSTRAT"),
           split_type = str_replace(split_type, "timesplit", "timeSTRAT"),
           classes = str_replace(classes, "original_classes", "Original classes"),
           classes = str_replace(classes, "reduced_classes", "Reduced classes"),
           behaviour_type = case_when(behaviour=="Forward aggressive" ~ "Mating behaviours",
                                      behaviour=="Static squat" ~ "Mating behaviours",
                                      behaviour=="Dynamic squat" ~ "Mating behaviours",
                                      behaviour=="Mounting male" ~ "Mating behaviours",
                                      behaviour=="Copulation attempt" ~ "Mating behaviours",
                                      behaviour=="Being mounted" ~ "Mating behaviours",
                                      classes=="Reduced classes" ~ "Combined behaviours",
                                      TRUE ~ "Non-mating behaviours"))

class_table_long <- class_table %>%
    pivot_longer(cols=c(prec,rec,fmeas),
                 names_to="metric")

class_table_wide <- class_table_long %>%
    pivot_wider(names_from=c(model_type, split_type),
                values_from=c(value))

# ---- Arrange table ----
class_table_wide <- class_table_wide %>%
    relocate('classes','behaviour', 'metric',
             'RF_randSTRAT', 'RF_timeSTRAT', 'RF_LSIO', 
             'HMM_timeSTRAT', 'HMM_LSIO',
             'NN_LSIO') %>%
    select(!classes) %>%
    arrange(behaviour)
```

```{r Format class table}
first_col_width <- 0.1
beh_col_width <- 4

class_table_flex <- 
    # ---- Basic formatting ----
    class_table_wide %>%
    filter(metric=="fmeas") %>%
    filter(behaviour!="Other") %>%
    select(!metric) %>%
    as_grouped_data(groups="behaviour_type") %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    set_header_labels(behaviour_type="") %>%
    bold(i = ~ !is.na(behaviour_type), part="body") %>%
    add_header_row(values = c("","RF", "HMM", "NN"),
                   colwidths = c(2, 3, 2, 1)) %>%
    set_header_labels(values = list(behaviour = "Behaviour",
                                    RF_randSTRAT = "randSTRAT",
                                    RF_timeSTRAT = "timeSTRAT",
                                    RF_LSIO = "LSIO",
                                    HMM_timeSTRAT = "timeSTRAT",
                                    HMM_LSIO="LSIO",
                                    NN_LSIO="LSIO")) %>%
    merge_at(i=1, j=1:7, part="body") %>% # Non-mating
    merge_at(i=8, j=1:7, part="body") %>% # Mating
    merge_at(i=15, j=1:7, part="body") %>% # Combined
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(port_width-first_col_width-beh_col_width)/(ncol(class_table_wide)-2), unit="cm") %>%
    width(j=1, first_col_width, "cm") %>%
    width(j=2, beh_col_width, "cm")

class_table_flex
```

```{r Save class metrics table}
save_as_docx(class_table_flex, align = "left",
             path = here(output_dir,"class_metrics_table.docx"))
```

# Correlation table (Table 4)

```{r Import correlation coefficients}
cors <- read.csv(here(input_dir, "correlation-coefficients.csv"))
```

```{r Summarise correlation coefficients}
cors <- cors %>%
    filter(classes=="original_classes"|behaviour=="courtship/territorial"|behaviour=="rest/vigilance")
```

```{r Order factors for correlation coefficients}
cors$behaviour <- rename_behaviours(cors$behaviour)
cors$behaviour <- order_behaviours(cors$behaviour)

cors$model_type2 <- ordered(cors$model_type, levels = c("RF", "HMM", "NN"))

cors$split_type2 <- ordered(cors$split_type, 
                           levels = c("randSTRAT", "timeSTRAT", "LSIO"))
```

```{r Correlation table with model columns}
# ---- Restructure table ----
cors_table <- cors %>%
    filter(transitions_in_test=="yes" | is.na(transitions_in_test)) %>%
    select(model_type, split_type, behaviour, cor_spearman, classes) %>%
    mutate(classes = str_replace(classes, "original_classes", "Original classes"),
           classes = str_replace(classes, "reduced_classes", "Reduced classes"),
           behaviour_type = case_when(behaviour=="Forward aggressive" ~ "Mating behaviours",
                                      behaviour=="Static squat" ~ "Mating behaviours",
                                      behaviour=="Dynamic squat" ~ "Mating behaviours",
                                      behaviour=="Mounting male" ~ "Mating behaviours",
                                      behaviour=="Copulation attempt" ~ "Mating behaviours",
                                      behaviour=="Being mounted" ~ "Mating behaviours",
                                      classes=="Reduced classes" ~ "Combined behaviours",
                                      TRUE ~ "Non-mating behaviours"))

cors_table_wide <- cors_table %>%
    #filter(split_type=="LSIO") %>%
    pivot_wider(names_from=c(model_type, split_type),
                values_from=cor_spearman)

# ---- Arrange table ----
cors_table_wide <- cors_table_wide %>%
    # relocate('classes','behaviour', 
    #          'RF_randSTRAT', 'RF_timeSTRAT', 'RF_LSIO', 
    #          'HMM_timeSTRAT', 'HMM_LSIO', 'NN_LSIO') %>%
    select(!classes) %>%
    arrange(behaviour)
```

```{r Select and rearrange correlation columns}
cors_table_tidy <- cors_table_wide %>%
    select(behaviour, behaviour_type,
           RF_randSTRAT, RF_timeSTRAT, RF_LSIO,
           HMM_timeSTRAT, HMM_LSIO,
           NN_LSIO) %>%
    filter(behaviour!="Other")
```

```{r Format correlation table table}
first_col_width <- 0.1
beh_col_width <- 4

cors_table_flex <- 
    # ---- Basic formatting ----
    cors_table_tidy %>%
    as_grouped_data(groups="behaviour_type") %>%
    flextable() %>%
    # merge_v(j = "behaviour", target = "behaviour", part = "body", combine = FALSE) %>%
    # valign(valign="top") %>%
    align(align = "left", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    set_header_labels(behaviour_type="") %>%
    bold(i = ~ !is.na(behaviour_type), part="body") %>%
    add_header_row(values = c("","RF", "HMM", "NN"),
                   colwidths = c(2, 3, 2, 1)) %>%
    set_header_labels(values = list(behaviour = "Behaviour",
                                    cor_spearman_RF_randSTRAT = "randSTRAT",
                                    cor_spearman_RF_timeSTRAT = "timeSTRAT",
                                    cor_spearman_RF_LSIO = "LSIO",
                                    cor_spearman_HMM_timeSTRAT = "timeSTRAT",
                                    cor_spearman_HMM_LSIO="LSIO",
                                    cor_spearman_NN_LSIO="LSIO")) %>%
    merge_at(i=1, j=1:7, part="body") %>% # Non-mating
    merge_at(i=8, j=1:7, part="body") %>% # Mating
    merge_at(i=15, j=1:7, part="body") %>% # Combined
    bold(part="header") %>%
    # ---- Set width ----
    width(width=(port_width-first_col_width-beh_col_width)/(ncol(cors_table_tidy)-2), unit="cm") %>%
    width(j=1, first_col_width, "cm") %>%
    width(j=2, beh_col_width, "cm")

cors_table_flex
```

```{r Save correlation table}
save_as_docx(cors_table_flex, path = here(output_dir,"correlation_table.docx"), 
             align = "left")
```

# Correlation plots (Figure 2)

```{r Subset to correlation values of interest and combine with F1 scores}
cors_table_sub <- cors_table_wide %>% 
    select(behaviour,
           behaviour_type,
           HMM_LSIO) %>%
    rename(r=HMM_LSIO)

class_table_sub <- class_table_wide %>% 
    filter(metric=="fmeas") %>%
    select(behaviour, HMM_LSIO) %>%
    rename(`F1-score`=HMM_LSIO) %>%
    mutate(`F1-score`=substr(`F1-score`,1,4))

cors_table_sub <- left_join(cors_table_sub, class_table_sub, by="behaviour")

cors_table_sub
```

## Get count data

```{r Import and format counts}
counts <- read.csv(here(input_dir, "true-vs-predicted_ind-beh-counts.csv"))

# Subset to focal birds and remove duplicates
counts <- counts %>% 
    filter(ruff_id %in% focal_birds) %>%
    filter(classes=="original_classes"|behaviour=="courtship/territorial"|behaviour=="rest/vigilance")

counts$behaviour <- rename_behaviours(counts$behaviour) 
```

```{r Filter counts}
counts_sub <- counts %>%
    filter(model_type=="HMM",
           split_type=="LSIO")
```

## Create plots

```{r Create base plot}
plot_cor <- function(behaviour_data, max, break_interval,
                     cors_table) {
    text_size = 10
    text_family = "Calibri"
    beh = behaviour_data$behaviour[1]
    f1 = subset(cors_table_sub, behaviour == beh)$`F1-score`
    r = subset(cors_table_sub, behaviour == beh)$r
    cor_plot <- ggplot(data=behaviour_data,
                       aes(x=perc_truth, y=perc_pred))+
        theme_classic()+
        geom_abline(slope=1, intercept=0, colour="grey", linetype="dotted")+
        coord_cartesian(xlim=c(0,max),
                        ylim=c(0,max))+
        scale_x_continuous(breaks=seq(0, max, by=break_interval),
                           expand=c(0.02,0.02))+
        scale_y_continuous(breaks=seq(0, max, by=break_interval),                    
                           expand=c(0.02,0.02))+
        xlab("\nTrue % time")+
        ylab("Estimated % time\n")+
        geom_point(shape=21, alpha=0.5, size=2, color="black", fill="#475a94") +
        theme(aspect.ratio=1,
              axis.text=element_text(size=text_size, family = text_family),
              #axis.title=element_text(size=text_size+1),
              axis.title=element_blank(),
              plot.title=element_text(size=text_size, family = text_family),
              plot.margin = unit(c(0.1, 0, 0.7, 0), "lines"))+
        ggtitle(label=paste0("  ", beh))+
        annotate(geom="text", 
                 label=paste0(" F1 = ", f1),
                 x = 0, y = Inf, hjust = 0, vjust = 1.1,
                 size=(text_size)/.pt, family=text_family)+
        annotate(geom="text",
                 label=paste0("\n", " r "),
                 x = 0, y = Inf, hjust = 0, vjust = 1.1,
                 size=(text_size)/.pt, family=text_family,
                 fontface = 'italic')+
        annotate(geom="text",
                 label=paste0("\n", "    = ", rnd(r)),
                 x = 0, y = Inf, hjust = 0, vjust = 1.1,
                 size=(text_size)/.pt, family=text_family)
        # geom_text(data = cors_table %>% filter(behaviour==behaviour),
        #           # size=3,
        #           # colour="darkgray",
        #           aes(label = PCC))#,
        #           # vjust = 0.5,
        #           # hjust = 0)
    return(cor_plot)
}

plot_cor(counts_sub, max=60, break_interval=10, cors_table_sub)
```

```{r Create correlation plots}
plot_a <- plot_cor(counts_sub %>% filter(behaviour=="Fly"),
                   max=6,
                   break_interval=1)#+
    # ylab("\n")+
    # xlab("\n")

plot_b <- plot_cor(counts_sub %>% filter(behaviour=="Forage or drink"),
                   max=25,
                   break_interval=5)#+
    # ylab("\n")+
    # xlab("\n")

plot_c <- plot_cor(counts_sub %>% filter(behaviour=="Preen"),
                   max=30,
                   break_interval=10)#+
    #     ylab("\n")+
    # xlab("\n")

plot_d <- plot_cor(counts_sub %>% filter(behaviour=="Rest"),
                   max=60,
                   break_interval=20)#+
    #     ylab("\n")+
    # xlab("\n")

plot_e <- plot_cor(counts_sub %>% filter(behaviour=="Vigilant"),
                   max=45,
                   break_interval=10)#+
    # ylab("\n")+
    # xlab("\n")

plot_f <- plot_cor(counts_sub %>% filter(behaviour=="Walk or run"),
                   max=25,
                   break_interval=5)#+
    # ylab("\n")+
    # xlab("\n")

plot_g <- plot_cor(counts_sub %>% filter(behaviour=="Forward aggressive"),
                   max=40,
                   break_interval=10)#+
    # xlab("\n")

plot_h <- plot_cor(counts_sub %>% filter(behaviour=="Dynamic squat"),
                   max=30,
                   break_interval=10)#+
    # ylab("\n")+
    # xlab("\n")

plot_i <- plot_cor(counts_sub %>% filter(behaviour=="Static squat"),
                   max=38,
                   break_interval=10)#+
    # ylab("\n")+
    # xlab("\n")

plot_j <- plot_cor(counts_sub %>% filter(behaviour=="Copulation attempt"),
                   max=3.2,
                   break_interval=1)#+
    # ylab("\n")+
    # xlab("\n")

plot_k <- plot_cor(counts_sub %>% filter(behaviour=="Mounting male"),
                   max=3.2,
                   break_interval=1)#+
    #     ylab("\n")+
    # xlab("\n")

plot_l <- plot_cor(counts_sub %>% filter(behaviour=="Being mounted"),
                   max=3.2,
                   break_interval=1)#+
    # ylab("\n")+
    # xlab("\n")

plot_m <- plot_cor(counts_sub %>% filter(behaviour=="Mating behaviours"),
                   max=65,
                   break_interval=20)+
    ggtitle(label=paste0("  ", "Mating behs (pooled)"))#+
    # ylab("\n")+
    # xlab("\n")

plot_n <- plot_cor(counts_sub %>% filter(behaviour=="Rest/vigilant"),
                   max=65,
                   break_interval=20)+
     ggtitle(label=paste0("  ", "Rest/vigilant (pooled)"))#+
    # ylab("\n")
```

```{r Combine plots, fig.width=5.06, fig.height=7.89}
grid_cor_plots2 <- grid.arrange(plot_a, plot_b, plot_c, 
                               plot_d, plot_e, plot_f,
                               plot_g, plot_h, plot_i,
                               plot_j, plot_k, plot_l,
                               plot_m, plot_n,
                               nrow = 5, ncol = 3)
```

```{r Save correlation plot}
ggsave(here(output_dir,"correlation_plots.png"), plot=grid_cor_plots2,
       width=13.8, height=21.5, units="cm", dpi=600)
```

# Confusion matrices (Table A5, Supplementary Tables)

```{r Gather predictions for confusion matrices}
rf_cv_preds <- read.csv(here(input_dir, "RF_best-CV-predictions_original-classes.csv"))
hmm_cv_preds <- read.csv(here(input_dir, "HMM_CV_predictions_original-classes.csv"))
nn_cv_preds <- read.csv(here(input_dir, "NN_CV_predictions_original-classes.csv"))

rf_loso_preds <- rf_cv_preds %>%
    filter(grepl("LSIO", model_category) & !grepl("ToutAll", model_category))

hmm_loso_preds <- hmm_cv_preds %>%
    filter(model_id=="hmm_LSIO")

nn_loso_preds <- nn_cv_preds
```

```{r Order factors}
rf_loso_preds$pred_beh <- rename_behaviours(rf_loso_preds$pred_beh)
rf_loso_preds$truth_beh <- rename_behaviours(rf_loso_preds$truth_beh)
rf_loso_preds$pred_beh <- order_behaviours(rf_loso_preds$pred_beh)
rf_loso_preds$truth_beh <- order_behaviours(rf_loso_preds$truth_beh)

hmm_loso_preds$pred_beh <- rename_behaviours(hmm_loso_preds$pred_beh)
hmm_loso_preds$truth_beh <- rename_behaviours(hmm_loso_preds$truth_beh)
hmm_loso_preds$pred_beh <- order_behaviours(hmm_loso_preds$pred_beh)
hmm_loso_preds$truth_beh <- order_behaviours(hmm_loso_preds$truth_beh)

nn_loso_preds$pred_beh <- rename_behaviours(nn_loso_preds$pred_beh)
nn_loso_preds$truth_beh <- rename_behaviours(nn_loso_preds$truth_beh)
nn_loso_preds$pred_beh <- order_behaviours(nn_loso_preds$pred_beh)
nn_loso_preds$truth_beh <- order_behaviours(nn_loso_preds$truth_beh)
```

```{r Create table properties}
confmat_beh_col_width <- 3
true_pos_face <- "bold"

format_confmat_ft <- function(x){
    x %>%
    rownames_to_column("True behaviour") %>%
    flextable() %>%
    align(align = "right", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    font(fontname="Calibri", part="all") %>%
    # ---- Rename and format headings ----
    add_header_row(values = c("", "Predicted behaviour"),
                   colwidths = c(1, ncol(x))) %>%
    align(i=1, align="center", part="header") %>%
    valign(i=2, valign="bottom", part="header") %>%
    bold(i=1, part="header") %>%
    bold(i=2, j=1, part="header") %>%
    set_header_labels(values = list(Fly = "FLY",
                                    `Forage or drink` = "FORAGE",
                                    Preen = "PREEN",
                                    Rest = "REST",
                                    Vigilant = "VIGIL",
                                    `Walk or run` = "WALK",
                                    Other = "OTHER",
                                    `Forward aggressive` = "AGGR",
                                    `Dynamic squat` = "DYN",
                                    `Static squat`="STAT",
                                    `Copulation attempt` = "COPUL",
                                    `Mounting male` = "MOUNT",
                                    `Being mounted` = "MOUNTED")) %>%
    # ---- Remove border from top left cell ----
    border(i=1,j=1, part="header",border = fp_border(color="white")) %>%
    border(i=1, part="header",border.top = fp_border(color="black")) %>%
    # ---- Set width ----
    width(width=(land_width-confmat_beh_col_width)/ncol(x), unit="cm") %>%
    width(j=1, confmat_beh_col_width, "cm") %>%
    # ---- Add italicisation to mating behaviours ----
    italic(i=2, j=9:14, part="header") %>%
    italic(i=8:13, j=1, part="body")%>%
    # ---- Format totals ----
    colformat_num(j = c("Total"), digits=0) %>%
    # ---- Add shading for true positives ----
    bold(i = 1, j = 2, part = "body") %>%
    bold(i = 2, j = 3, part = "body") %>%
    bold(i = 3, j = 4, part = "body") %>%
    bold(i = 4, j = 5, part = "body") %>%
    bold(i = 5, j = 6, part = "body") %>%
    bold(i = 6, j = 7, part = "body") %>%
    bold(i = 7, j = 8, part = "body") %>%
    bold(i = 8, j = 9, part = "body") %>%
    bold(i = 9, j = 10, part = "body") %>%
    bold(i = 10, j = 11, part = "body") %>%
    bold(i = 11, j = 12, part = "body") %>%
    bold(i = 12, j = 13, part = "body") %>%
    bold(i = 13, j = 14, part = "body")
}
```

```{r Create RF confmat table, width=50}
# ---- Create confusion matrix ----
# NOTE: Rows (first) are truth, Columns (second) are predictions
rf_confmat <- as.data.frame.matrix(table(rf_loso_preds$truth_beh, rf_loso_preds$pred_beh))

# ---- Refine confusion matrix ----
rf_confmat <- rf_confmat %>%
    filter(row.names(rf_confmat)!="Mating behaviours" &
               row.names(rf_confmat)!="Rest/vigilant") %>%
    select(!c("Mating behaviours","Rest/vigilant"))

rf_confmat <- bind_rows(rf_confmat,
                        rf_confmat %>% summarise_all(sum)) %>%
    mutate(Total=rowSums(.))

row.names(rf_confmat)[nrow(rf_confmat)] <- "Total"

rf_confmat_flex <- format_confmat_ft(rf_confmat)

rf_confmat_flex
```

```{r Create hmm confmat table, width=50}
# ---- Create confusion matrix ----
# NOTE: Rows (first) are truth, Columns (second) are predictions
hmm_confmat <- as.data.frame.matrix(table(hmm_loso_preds$truth_beh, hmm_loso_preds$pred_beh))

# ---- Refine confusion matrix ----
hmm_confmat <- hmm_confmat %>%
    filter(row.names(hmm_confmat)!="Mating behaviours" &
               row.names(hmm_confmat)!="Rest/vigilant") %>%
    select(!c("Mating behaviours","Rest/vigilant")) 

hmm_confmat <- bind_rows(hmm_confmat, 
                        hmm_confmat %>% summarise_all(sum)) %>%
    mutate(Total=rowSums(.))

row.names(hmm_confmat)[nrow(hmm_confmat)] <- "Total"

hmm_confmat_flex <- format_confmat_ft(hmm_confmat)

hmm_confmat_flex
```

```{r Create NN confmat table, width=50}
# ---- Create confusion matrix ----
# NOTE: Rows (first) are truth, Columns (second) are predictions
nn_confmat <- as.data.frame.matrix(table(nn_loso_preds$truth_beh, nn_loso_preds$pred_beh))

# ---- Refine confusion matrix ----
nn_confmat <- nn_confmat %>%
    filter(row.names(nn_confmat)!="Mating behaviours" &
               row.names(nn_confmat)!="Rest/vigilant") %>%
    select(!c("Mating behaviours","Rest/vigilant"))

nn_confmat <- bind_rows(nn_confmat,
                        nn_confmat %>% summarise_all(sum)) %>%
    mutate(Total=rowSums(.))

row.names(nn_confmat)[nrow(nn_confmat)] <- "Total"

nn_confmat_flex <- format_confmat_ft(nn_confmat)

nn_confmat_flex
```

```{r Save confusion matrix table}
save_as_docx(`HMM confusion matrix` = hmm_confmat_flex, 
             `RF confusion matrix` = rf_confmat_flex,
             `NN confusion matrix` = nn_confmat_flex,
             path = here(output_dir, "confusion_matrices.docx"), 
             align = "left")
```

# Upsampling and transition table (Table 5)

```{r Import cross-validation results}
rf_all_cv <- fread(here(input_dir, "RF_all-CV-macro-metrics_original-classes.csv"))
```

```{r Filter rf cv results}
# LSIO models with same hyperparameters as top LSIO
rf_sub_cv <- rf_all_cv %>% 
    filter(min_n==1 & mtry==2) %>%
    filter(grepl("LSIO",model_id))

# Either same upsampling OR same transition treatment
rf_sub_cv <- rf_sub_cv %>%
    filter(grepl("sm10", model_id)|grepl("Tin", model_id))

# Format with mean and sd
up_tr_table <- rf_sub_cv %>%
    mutate(all_prec=format_mean_sd(precision, precision_sd),
           all_recall=format_mean_sd(recall, recall_sd),
           all_fmeas=format_mean_sd(f_meas, f_meas_sd),
           mat_prec=format_mean_sd(precision_mat, precision_sd_mat),
           mat_recall=format_mean_sd(recall_mat, recall_sd_mat),
           mat_fmeas=format_mean_sd(f_meas_mat, f_meas_sd_mat)) %>%
    select(model_id, all_prec:mat_fmeas)

# Replicate the best model (useful further below) (using row number here)
up_tr_table <- rbind(up_tr_table, up_tr_table[3,])
```

```{r Reorder table}
up_tr_table$order <- c(1,2,4,5,7,3,6)
up_tr_table <- up_tr_table %>% arrange(order)
up_tr_table$category <- c(rep("Upsampling",4),rep("Transitions",3))
```

```{r Format upsampling transition table as flextable}
# ~~~~~ FORMAT UPSAMPLE/TRANSITION TABLE ---------------------------------------
first_col_width <- 0.1

up_tr_table_flex <- 
    # ---- Basic formatting ----
    up_tr_table %>%
    select(-order) %>%
    as_grouped_data(groups="category") %>%
    flextable() %>%
    align(align = "left", part="all") %>%
    line_spacing(space = 1.25, part = "all") %>%
    add_header_row(values = c("","All behaviours", "Mating behaviours"),
                   colwidths = c(2, 3, 3)) %>%
    bold(part="header") %>%
    font(fontname="Calibri", part="all") %>%
    width(width=(port_width-first_col_width)/ncol(up_tr_table), unit="cm") %>%
    width(j=1, first_col_width, "cm") %>%
    # ---- Rename and format headings ----
    set_header_labels(category="") %>%
    bold(i = ~ !is.na(category), part="body") %>%
    set_header_labels(values = list(model_id = "Model",
                                    all_prec = "Precision",
                                    all_recall = "Sensitivity",
                                    all_fmeas = "F1-score",
                                    mat_prec = "Precision",
                                    mat_recall = "Sensitivity",
                                    mat_fmeas = "F1-score")) %>%
    merge_at(i=1, j=1:8, part="body") %>% # Upsampling
    merge_at(i=6, j=1:8, part="body")  # Inclusion of transitions

up_tr_table_flex
```

```{r Save upsampling and transitions table}
save_as_docx(up_tr_table_flex, path = here(output_dir,"upsampling_transitions_table.docx"), 
             align = "left")
```